---
title: 'Homework 9'
author: 
  - Roy
  - Xuan Wen
date: "4/6/2023"
output: html_document
---

Team members:

- Roy Khoo
- Xuan Wen Loo

Git repo: https://github.com/roy-kcj/practice4

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
# Read data
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

# Part one: Data tidying

```{r}
av_deaths = av %>%
  gather(key = "Time", value = "Death", starts_with("Death")) %>%
  mutate(Time = parse_number(Time),
         Death = ifelse(Death == "NO", "no", ifelse(Death == "YES", "yes", "")),
         Death = factor(Death)) %>%
  select(URL, 
         Name.Alias, 
         Appearances, Current., 
         Gender, 
         Probationary.Introl, 
         Year, 
         Years.since.joining, 
         Honorary,
         Time,
         Death)

av_return = av %>%
  gather(key = "Tim", value = "Return", starts_with("Return")) %>%
  mutate(Time = parse_number(Tim),
         Return = ifelse(Return == "NO", "no", ifelse(Return == "YES", "yes", "")),
         Return = factor(Return)) %>%
  select(URL, Name.Alias, 
         Appearances, Current., 
         Gender, 
         Probationary.Introl, 
         Year, 
         Years.since.joining, 
         Honorary,
         Time,
         Return)
```

# Part two: Validation

Roy:

Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team. That’s about 40 percent of all people who have ever signed on to the team. 

```{r}
num_death = av_deaths %>% 
  filter(Death == "yes") %>% 
  distinct(URL) %>%
  nrow()

# 69 observations as pointed out by the article

total = av %>% nrow()

pct_death = num_death / total * 100

sprintf("Out of 173 listed Avengers, %d has died at least once since joining and that's %.2f%% of all people who joined", num_death, pct_death)

# 39.88 is similar enough to 40 to round up
```

Xuan Wen:

There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death.

```{r}
death1 = av_deaths %>% 
  filter(Death == "yes", Time == 1) %>% 
  distinct(URL) %>%
  nrow()

return1 = av_return %>% 
  filter(Return == "yes", Time == 1) %>% 
  distinct(URL) %>%
  nrow()

chance.return1 = return1 / death1 *100

cat(death1, "Avengers have died at least once and", return1, "returned from the first death. That's", chance.return1, "% of returning from their first stint.\n\n")
# 66.66667% is similar to 2/3

death2.3 = av_deaths %>% 
  filter(Death == "yes", Time == 2 | Time == 3) %>% 
  distinct(URL) %>%
  nrow()

return2.3 = av_return %>% 
  filter(Return == "yes", Time == 2 | Time == 3) %>% 
  distinct(URL) %>%
  nrow()

chance.return2.3 = return2.3 / death2.3 * 100

cat(death2.3, "Avengers have died 2 or 3 times and", return2.3, "returned. That's", chance.return2.3, "% of returning from the 2nd or 3rd death.")
```
